---
title: "SRT411_ASSIGNMENT_0"
author: "Isham Sidhu"
date: "February 19, 2019"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
###INTRODUCTION:###
"R is a powerful language and environment for statistical computing and graphics.  It is a public domain (a so called "GNU") project which is similar to  the  commercial  S  language  and  environment which  was  developed  at  Bell  Laboratories  (for-merly AT&T, now Lucent Technologies) by John Chambers and colleagues.  R can be considered as a different implementation of S, and is much used in as an educational language and research tool. It is quite similar to other programming packages such as MatLab (not freeware), but moreuser-friendly than programming languages such asC++ or Fortran. 
In this assignment we have to complete the ToDo list from the document in this link and will be reading and understanding content from this web file. After completing these ToDo in R markdown we have to convert this .Rmd file into the PDF using Knit. Further, we wil be making an account in the GitHub and make a repository that includes the .Rmd file and PDF file of the R code and output and one read me file which will explain the assignment"




1) Compute the difference between 2014 and the year  you  started  at  this  university  and  divide this by the difference between 2014 and the year you  were  born.   Multiply  this  with  100  to  get the  percentage  of  your  life  you  have  spent  at this university.  Use brackets if you need them.

```{r}
((2018-2014)/(2014-1999))*100
```

2)  Repeat  the  previous  ToDo,  but  with  several steps  in  between.   You  can  give  the  variables any name you want, but the name has to start with a letter.


```{r}
years_university = 2018-2014 
years_life = 2014-1999 
university_to_life = (years_university/years_life)*100 
university_to_life
```


3) Compute the sum of 4, 5, 8 and 11 by first combining  them  into  a  vector  and  then  using  the function sum.


```{r}
b=c(4,5,8,11) 
sum(b)
```


4) Plot 100 normal random numbers.


```{r}
z=rnorm(100) 
plot(z)
```

 
5) Find help for the sqrt function.


```{r}
help("sqrt")
```


6) Make a file called firstscript.R containing R-code  that  generates  100  random  numbers  and plots them, and run this script several times.


```{r}
plot(rnorm(100))
```


7) Put  the  numbers  31  to  60  in  a  vector  named P and in a matrix with 6 rows and 5 columns named Q. Tip:  use the function seq.  Look at the different ways scalars, vectors and matrices are denoted in the workspace window.


```{r}
P = seq(from=31, to=60, by=1)
Q= matrix(P,ncol = 5, nrow = 6)
P
```
```{r}
Q
```


8) Make  a  script  file  which  constructs  three  random  normal  vectors  of  length  100.   Call  these vectors x1, x2 and x3. Make a data frame called t with three columns (called a,b and c) containing  respectively x1, x1+x2 and x1+x2+x3.Call the following functions for this data frame: plot(t) and sd(t).  Can you understand the results?  Rerun this script a few times.


```{r}
x1 = seq(1:100) 
x2 = seq(1:100) 
x3 = seq(1:100) 
t= data.frame(a=x1, b=x1+x2, c=x1+x2+x3) 
plot(t)
```


9) Add these lines to the script file of the previous section.  Try to find out, either by experimenting or by using the help, what the meaning is of rgb, the last argument of rgb,lwd, pch, cex.

>Rgb Generates colour 

>lwd = line width 

>Pch= plotting 'character', i.e., symbol to use. This can either be a single character or an integer code for one of    a set of graphics symbols. The full set of S symbols is available with pch = 0:18. 

>Cex: For other text symbols, cex = 1 corresponds to the default fontsize of the device, often specified by an         argument pointsize. For pch in 0:25 the default size is about 75% of the character height (see par("cin")) 

```{r}
plot(t$a, type="l", ylim=range(t),lwd=3, col=rgb(1,1,0,0.7))
lines(t$b, type="s", lwd=2,col=rgb(0.3,0.4,0.3,0.9))
points(t$c, pch=22, cex=2,col=rgb(0,0,1,0.3))
```

10) Make  a  file  called tst1.txt in  Notepad  from the  example  in  Figure  4  and  store  it  in  your working directory.  Write a script to read it, to multiply the column called g by 5 and to store it as tst2.txt.


```{r}
d=data.frame(a=c(1,2,4,8,16,32),g=c(2,4,8,16,32,64),x=c(3,6,12,24,48,96)) 
write.table(d,file="tst1.txt",row.names=FALSE) 
t = read.table("tst1.txt",header = TRUE) 
t2=data.frame(a=t$a, g=t$g*5, x=t$x) 
write.table(t2,file="tst2.txt",row.names=FALSE)
```


11) Compute the mean of the square root of a vector  of  100  random  numbers.   What  happens?


```{r}
mean(sqrt(rnorm(100)+10))
```
"mean(sqrt(rnorm(100)))" gives an error message as rnorm give negative random numbers too. 



12) Make  a  graph  with  on  the  x-axis:  today,  Sinterklaas  2014  and  your  next  birthday  and  on the y-axis the number of presents you expect on each of these days.  Tip:  make two vectors first.


```{r}
date=strptime(c("20190220","20141205","20200110"),format = "%Y%m%d") 
presents=c(0,1,5) 
plot(date,presents) 
```


13) Make a vector from 1 to 100.  Make a for-loop which runs through the whole vector.  Multiply the elements which are smaller than 5 and larger than 90 with 10 and the other elements with 0.1.


```{r}
vector=seq(1:100) 
s=c() 
for(i in 1:100) 
 { 
     if(vector[i]<5) 
 
         { s[i]=vector[i]*5 } 
 
     else if(vector[i]>90) 
 
         { s[i]=vector[i]*10; } 
 
     else 
 
         { s[i]=vector[i]*0.1; } 
  } 
 s 
```

14) Write  a  function  for  the  previous  ToDo,  so that   you   can   feed   it   any   vector   you   like(as  argument).Use  a  for-loop  in  the  function  to  do  the  computation  with  each  element.    Use  the  standard  R  function length in   the   specification   of   the   counter.a)


```{r}
z=1:100 
fun = function(arg1) 
 { 
     l = length(arg1) 
     for(i in 1:l) 
     {  
 
         if (arg1[i] < 5) 
 
             { arg1[i] = arg1[i] * 5 } 
 
         else if (arg1[i] > 90) 
 
             { arg1[i] = arg1[i] * 10}  
 
         else 
 
             { arg1[i] = arg1[i] * 0.1 } 
 
     } 
 return (arg1) 
 } 
 fun(z) 
```